A Quantitative Genetic Linked Selection Model fit to the Negative Selection Process in Humans

### 2023-07-05 manuscript/manuscript.tex
ms changes

We propose that one mechanism for this non-identifiability is that mutations
extremely harmful to fitness are difficult to distinguish from a smaller
effective population. Intuitively, as $s$ increases, the average persistence
time (in generations) of the mutation in the population goes below one
generation, i.e. death of the carrier of the \emph{de novo} mutation. In this
hypothetical limit, each occurrence of a very harmful mutation is like an
instantaneous death and thus acts like a genome-wide reduced effective
population size. In our model, this genome-wide effective population size is
parameterized by $\pi_0$; thus, one line of evidence for this effect is a
correlation between estimates of $\pi_0$ and average selection $\bar{s}$.
Looking at the CADD 6\% models fit across the default and constrained DFE
grids, we find this is the case. Across every population, average selection was
about 7 times larger using the default grid, and $\pi_0$ was $5.6\%$ larger.
However, there was no similar consistent change in mutation rate estimates
among populations. Overall, this suggests that the linked selection signal
alone carries little information that can differentiate very strong selection
from a slightly smaller drift-effective population size.

Despite the strong-selection non-identifiability issue, we still were curious
how stable these estimates are across samples from reference populations. In
particular, we focused on the Feature Priority model, which fits the data less
well but is more interpretable since its annotation is based on gene features.
Overall, estimates of the distribution of selection coefficients are stable
across populations and had narrow uncertainty as estimated by a block jackknife
procedure (Figure \ref{figure-3}B). Coding sequences show the strong-selection
non-identifiability issue, and an uncertain signal of a neutral mode in the
DFE. This would be expected given the very different levels of selective
constraint synonymous and non-synonymous sites in coding sequences are under.
The PhastCons class of this Feature Priority model, which contains non-coding
highly conserved elements (XXX), is under strong selective constraint as
expected.

@@

\begin{figure}[htbp] 
    \centering
    \includegraphics[width=\textwidth]{figures/figure_4.pdf} 
    \caption{}
  \label{fig:figure-4}
\end{figure}

@@

\subsection*{Model Predictions and the Substitution Process}

>>>>

We propose that one mechanism for this non-identifiability is that mutations
extremely harmful to fitness are difficult to distinguish from a smaller
effective population. Intuitively, as $s$ increases, the average persistence
time (in generations) of the mutation in the population goes below one
generation, i.e. death of the carrier of the \emph{de novo} mutation. In this
hypothetical limit, each occurrence of a very harmful mutation is like an
instantaneous death and thus acts like a genome-wide reduced effective
population size. In our model, this genome-wide effective population size is
parameterized by $\pi_0$; thus, one line of evidence for this effect is a
correlation between estimates of $\pi_0$ and average selection $\bar{s}$.
Looking at the CADD 6\% models fit across the default and constrained DFE
grids, we find this is the case. Across every population, average selection was
about 7 times larger using the default grid, and $\pi_0$ was $5.6\%$ larger.
There was no similar consistent change in mutation rate estimates among
populations. In the CADD 6\% model, genome-wide average reduction factor
$\bar{B}$ was $\approx 6.1\%$ lower in the default versus constrained grid.
Overall, this suggests that the linked selection signal alone carries little
information that can differentiate very strong selection from a slightly
smaller drift-effective population size.

Despite the strong-selection non-identifiability issue, we still were curious
how stable these estimates are across samples from reference populations. In
particular, we focused on the Feature Priority model, which fits the data less
well but is more interpretable since its annotation is based on gene features.
Overall, estimates of the distribution of selection coefficients are stable
across populations and had narrow uncertainty as estimated by a block jackknife
procedure (Figure \ref{figure-3}B). Coding sequences show the strong-selection
non-identifiability issue, and an uncertain signal of a neutral mode in the
DFE. This would be expected given the very different levels of selective
constraint synonymous and non-synonymous sites in coding sequences are under.
The PhastCons class of this Feature Priority model, which contains highly
conserved non-coding elements (XXX), is under strong selective constraint as
expected. Interestingly, in our PhastCons Priority annotation model (Figure
\ref{fig:figure-3}A bottom row), the DFE estimate for coding sequences
excluding PhastCons regions is neutral. This is expected; the selection signal
in coding regions is absorbed by the PhastCons feature.

@@

\begin{figure}[htbp] 
    \centering
    \includegraphics[width=\textwidth]{figures/figure_4.pdf} 
    \caption{(A) }
  \label{fig:figure-4}
\end{figure}

Early genome-wide models of background selection closely fit megabase-scale
patterns in sequence divergence among five primates
\parencite{McVicker2009-ax}. However, the mutation rate estimates from this
model were unreasonably high compared to pedigree-based estimates (XXX). This
mutation rate discordance has led to the hypotheses that positive selection
could explain some fraction of the genome-wide linked selection signal (e.g.
\cite{Enard2014-kz}); however, evidence suggests that the role of classic
sweeps in humans is minuscule
\parencite{Pickrell2009-tt,Hernandez2011-gs,Murphy2022-sj}. Furthermore, Murphy
et al. find the classic BGS model fit using the CADD 6\% track leads to
mutation rate estimates that are consistent with pedigree-based rates. Given
that our method is based on a different theoretic approach, we curious whether
it would lead to estimates consistent with pedigree-based mutation rates.

Overall, across our CADD-based models and all populations, our mutation rate
estimates are roughly concordant with pedigree-based estimates (Figure
\ref{fig:figure-4}A). Our full-track CADD 6\% model estimates a mutation rate
of $\mu = 1.46 \times 10^{-8}$ for YRI, $1.53 \times 10^{-8}$ for CEU, and $1.5
\times 10^{-8}$ for CHB reference samples (Supplementary Table
\ref{supp:tbl-r2}). Mutation rate estimates for our feature-based annotation
models are unrealistically higher. We believe this could be caused by
per-basepair heterogeneity in selective effects within each of these features.
For example, mutations in coding sequences are known to have highly
heterogeneous effects depending on whether the result is synonymous,
non-synonymous, or mis-sense. If some fraction of the impact of negative
selection is due to strong selection at a few sites, sparser annotation models
like these cannot isolate these effects like CADD-based models. Instead, XXXHERE

reflects a general problem
with these more interpretable, but sparser feature-based tracks: heterogeneity
in selective effects

However, we noticed that estimates of mutation rate is very sensitive to the
choice of annotation model. Between CADD 6\% and 6\% models, the

\subsection*{Tests of Parameter Inference using Forward Simulations}

Given

@@
